Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-09-01 paper/paper.tex
update body + abstract

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
\sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the results from the multiphase model are broadly
consistent with observations of other CDGs. 
If \tol\ is dispersion dominated, the clump velocity dispersion
implies a dynamical mass of $2\times 10^{9}$ M$_{\odot}$, ten times
its baryonic mass, which can be explained if \tol\ is hosted by a dark
matter halo $6\times 10^{11}$ M$_{\odot}$. 
If future kinematic maps of \tol\ confirm the velocities suggested by
the multiphase model, it would provide additional support to expect
such kinematic state in primordial \lya emitting galaxies.   
\end{abstract}

@@

Observations of other Compact Dwarf Galaxies (CDGs) also show 
that the kinematics can renge from pure rotation and low velocity
dispersion to high velocity dispersion without rotaion.
\citep{2015A&A...577A..21C,2017A&A...600A.125C}.

The models we use correspond to simplified geometrical configurations.
This allows us to perform a deep exploration of parameter space and
gain some physical insight into \tol's kinematic properties.
In this paper we show, for the first time in the literature, that \tol's
\lya profile can be explicitly modeled by either of these two
models. 
The hybrid case of a rotating multiphase configuration
(as observed in some of the galaxies in \cite{2015A&A...577A..21C}) is
the subject of ongoing research (Remolina-Gutierrez et al. in prep.).  as
it cannot be simulated with existing models.

@@

These results disfavor the high rotational velocity of $V_{m}\sim350$
\kms\  that we find in the pure rotation model. 
On the other hand the results from the multiphase model yield a
velocity dispersion $\sigma=$\sigmaclump, quite close to observed
values in other CDG, favoring the multiphase model.

In this case we can  estimate a value for the dynamical
mass using the constraints for the velocity dispersion, $\sigma$,  in
a spherical system localized in a region of size $r$

@@

Following the methodology of \citet{2011ApJ...726..108T} we estimate 
that a dark matter halo of at $\sim 6\times 10^{11}$ M$_{\odot}$ in
virial mass can explain this dynamical mass.
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P,2016ApJ...832..169P}.  
Figure \ref{fig:mass} shows the enclosed mass as a function of radius
for different dark matter halos and the estimates for \tol.

New IFU observations are needed to completeley confirm \tol's.
This could be done with the Multi Unit Spectroscopic Explorer (MUSE) 
\citep{2014Msngr.157...13B}, the Gemini Multi-Object Spectrographs (GMOS) 
\citep{2004PASP..116..425H} or the Fibre Large Array Multi Element
Spectrograph (\citep{2002Msngr.110....1P})

>>>>

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
\sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the results from the multiphase model are broadly
consistent with observations of other CDGs. 
If \tol\ is dispersion dominated, the clump velocity dispersion
implies a dynamical mass of $2\times 10^{9}$ M$_{\odot}$, ten times
its baryonic mass, which can be explained if \tol\ is hosted by a dark
matter halo $6\times 10^{11}$ M$_{\odot}$. 
If future kinematic maps of \tol\ confirm the velocities suggested by
the multiphase model, it would provide additional support to expect
such kinematic state in primordial \lya emitting galaxies which in
turn can help to constrain star formation and feedback in the first
generation of galaxies.
\end{abstract}

@@

Furthermore, observations of other Compact Dwarf Galaxies (CDGs) also show 
that the kinematics range from pure rotation and low velocity
dispersion to high velocity dispersion without rotaion
\citep{2015A&A...577A..21C,2017A&A...600A.125C}.

The models we use correspond to simplified geometrical configurations.
This allows us to perform a deep exploration of parameter space and
gain some physical insight into \tol's kinematic properties.
In this paper we show, for the first time in the literature, that \tol's
\lya profile can be explicitly modeled by either of these two
models. 
The hybrid case of a rotating multiphase configuration
(as observed in some of the galaxies in \cite{2015A&A...577A..21C}) is
the subject of ongoing research (Remolina-Gutierrez et al. in prep.)  as
it cannot be simulated with existing models.

@@

These results disfavor the high rotational velocity of $V_{m}\sim350$
\kms\  that we find in the pure rotation model. 
On the other hand the results from the multiphase model yield a
velocity dispersion $\sigma=$\sigmaclump, quite close to observed
values in other CDG, favoring the multiphase model.
In this case we can  estimate a value for the dynamical
mass using the constraints for the velocity dispersion, $\sigma$,  in
a spherical system localized in a region of size $r$

@@

Following the methodology of \citet{2011ApJ...726..108T} we estimate 
that a dark matter halo of at $\sim 6\times 10^{11}$ M$_{\odot}$ in
virial mass can explain this dynamical mass.
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the median mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P,2016ApJ...832..169P}.  
Figure \ref{fig:mass} shows the enclosed mass as a function of radius
for different dark matter halos together with the \tol's dynamical
mass estimate.

New IFU observations are needed to completeley confirm \tol's.
This could be done with the Multi Unit Spectroscopic Explorer (MUSE) 
\citep{2014Msngr.157...13B}, the Gemini Multi-Object Spectrographs (GMOS) 
\citep{2004PASP..116..425H} or the Fibre Large Array Multi Element
Spectrograph (\citep{2002Msngr.110....1P}) as they have the spatial
resolution and field of view required to perform the kind of study
presented by \citep{2015A&A...577A..21C} on CDGs or \citep{Herenz16}
in nearby \lya\ emitting galaxies.
